---
title: "R Notebook for ABC model classification using currently implemented methods in the abc package"
output:
  html_document: default
  html_notebook: default
---

The script uses xgboost for model classiifcation in ABC. We consider the simulations provided in the vignette of the R abc package. These simulations are used to discriminate between different demographic models and are contained in the *human* data.

```{r}
require(abc)
require(abc.data)
data(human)
```


Distribution of the sums stats for the different models
```{r}
par(mfcol = c(1,3))
boxplot(stat.3pops.sim[,"pi"]~models, main="Mean nucleotide diversity")
boxplot(stat.3pops.sim[,"TajD.m"]~models, main="Mean Tajima's D")
boxplot(stat.3pops.sim[,"TajD.v"]~models, main="Var in Tajima's D")
```
Summary stats for human resequencing data
```{r}
stat.voight
```
 Running neural nets for the Voight et al. data
```{r}
modsel.it <- postpr(stat.voight["italian",], models, stat.3pops.sim, tol=.05, method="neuralnet",trace=F)
print(summary(modsel.it,print=F)[2]$neuralnet$Prob)
```
Evaluate model classification with 3 different methods currently implemented in the package
```{r,cache=TRUE}

#modsel.rejection <- cv4postpr(models,stat.3pops.sim, tol=.05,nval=500,method="rejection",trace=F)

#modsel.logistic <- cv4postpr(models,stat.3pops.sim, tol=.05,nval=500,method="mnlogistic",trace=F)

#modsel.nnet <- cv4postpr(models,stat.3pops.sim, tol=.05,nval=500,method="neuralnet",trace=F)

#save(modsel.rejection,modsel.logistic,modsel.nnet,file="cv.Rdata")
load(file="cv.Rdata")
```

```{r}
cat("misclassif rate\n","rejection   ",1-sum(diag(table(modsel.rejection$true,modsel.rejection$estim[[1]])))/1500,"\n","logistic ",1-sum(diag(table(modsel.logistic$true,modsel.logistic$estim[[1]])))/1500,"\n","nnet ",1-sum(diag(table(modsel.nnet$true,modsel.nnet$estim[[1]])))/1500)
```


```{r}
require(caret)
#plot for calibration
ff<-function(modsel,char,aux)
{
  if(aux==1)
{
bottleneck<-factor(models[modsel$cvsamples]!="bott")
probs<-modsel$model.probs[[1]][,"bott"]
calPlotData <- calibration(bottleneck~ probs)
plot(calPlotData, main=char,type = "l",xlab="prob bottleneck", auto.key = list(columns = 3,
                                          lines = TRUE,
                                          points = FALSE))
}
else if (aux==2)
{
bottleneck<-factor(models[modsel$cvsamples]!="exp")
probs<-modsel$model.probs[[1]][,"exp"]
calPlotData <- calibration(bottleneck~ probs)
plot(calPlotData, type = "l",main=char,xlab="prob expansion", auto.key = list(columns = 3,
                                          lines = TRUE,
                                          points = FALSE),add=T,col="green")
}
else
{
bottleneck<-factor(models[modsel$cvsamples]!="const")
probs<-modsel$model.probs[[1]][,"const"]
calPlotData <- calibration(bottleneck~ probs)
plot(calPlotData, type = "l",main=char,xlab="prob const", auto.key = list(columns = 3,
                                          lines = TRUE,
                                          points = FALSE),add=T,col="green")
}
}


ff(modsel=modsel.rejection,char="rejection",1)
ff(modsel=modsel.rejection,char="rejection",2)
ff(modsel=modsel.rejection,char="rejection",3)

ff(modsel=modsel.logistic,char="logistic",1)
ff(modsel=modsel.logistic,char="logistic",2)
ff(modsel=modsel.logistic,char="logistic",3)

ff(modsel=modsel.nnet,char="nnet",1)
ff(modsel=modsel.nnet,char="nnet",2)
ff(modsel=modsel.nnet,char="nnet",3)

```


